---
title: "PoNet: Invite Only – ALAAM Estimation"
subtitle: >
  This file estimates ALAAMs as reported in the Journal of Criminal Justice article:  
  *“Invite Only: The Prevalence of Subgroups within a Police Department”*  
  ([DOI](https://doi.org/10.1016/j.jcrimjus.2024.102347))
author: "Police Networks Project"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE)
```

# Overview
This report formats the input data and runs ALAAMs to assess network differences across officers invited/not invited to subgroups within a police network. Due to computational intensity, `eval=FALSE` prevents execution when knitting.

---

# Import Data
Data is produced via `ponet_inviteonly-01.Rmd`. Files must be pre-saved in `ALAAM/`.

Import data
```{r}
setwd("C:/Users/ouell/Desktop/PONET/project") # set to your working directory

readRDS(file = "ALAAM/net.rds")
readRDS(file = "ALAAM/outcome.rds")
readRDS(file = "ALAAM/attributes.rds")
```

# Format Network and Outcome
Convert the network to matrix form and extract the binary outcome. 
```{r}
n2 <- nrow(net2)
adj2 <- as.matrix(net2) # convert from data frame to matrix
invite <- outcome
```

# Load ALAAM Routines
Requires `MultivarALAAMalt.R` from [GitHub ALAAM repository](https://github.com/johankoskinen/ALAAM).

```{r}
source("MultivarALAAMalt.R")
```

---

# Generate Network Structural Covariates
Calculates 1- to 3-star structures, reciprocity, and transitivity.
```{r structur}
out.degree     <- matrix( rowSums(adj2), n2, 1)             # number of ties sent
in.degree      <- matrix( colSums(adj2) , n2, 1 )           # number of ties received
rec.ties       <- matrix( rowSums(adj2 * t(adj2) ), n2 , 1) # number of ties that are mutual
in.two.star    <- matrix( choose(in.degree,2),n2,1)         # in-stars reflecting dispersion in popularity
out.two.star   <- matrix( choose(out.degree,2),n2,1)        # out-stars reflecting dispersion in activity
mix.two.star   <- in.degree*out.degree - rec.ties           # correlation between indegree and outdegree
in.three.star  <- matrix( choose(in.degree,3),n2,1)         # further measure of in-degree heterogeneity
out.three.star <- matrix( choose(out.degree,3),n2,1)        # further measure of out-degree heterogeneity
triangles      <- rowSums( adj2* (adj2 %*% t(adj2) )  )     # embedded in transitive triads
```

---

# Format Attributes and Combine with Network Stats
Ensure numeric formatting, handle missingness, and bind all covariates.
```{r formatcovs}
# convert to numeric vars
attributes$male       <- as.numeric(as.character(attributes$male))
attributes$white      <- as.numeric(as.character(attributes$white))
attributes$hisp       <- as.numeric(as.character(attributes$hisp))
attributes$officer    <- as.numeric(as.character(attributes$officer))
attributes$street1    <- as.numeric(as.character(attributes$street1))
attributes$proactive1 <- as.numeric(attributes$proactive1)
attributes$nyears     <- as.numeric(attributes$nyears)
attributes$gun        <- as.numeric(attributes$gun)
attributes$taser      <- as.numeric(attributes$taser)
attributes$force      <- as.numeric(attributes$force)
attributes$nops2      <- as.numeric(attributes$nops2)

# Function to calculate the mode
get_mode <- function(x) {
  uniq_vals <- unique(x)
  uniq_vals[which.max(tabulate(match(x, uniq_vals)))]
}

# replace missing w/ mode
attributes <- attributes %>%
  mutate(across(everything(), ~replace_na(., get_mode(.))))

covs2 <- cbind(attributes[,1], 
               attributes[,2], 
               attributes[,3], 
               attributes[,4], 
               attributes[,5], 
               attributes[,6], 
               attributes[,7], 
               attributes[,8], 
               attributes[,9], 
               attributes[,10], 
               attributes[,11], 
               attributes[,12], 
               out.degree, 
               in.degree,
               rec.ties,
               in.two.star,
               out.two.star,
               mix.two.star,
               in.three.star,
               out.three.star,
               triangles)

# set colnames
colnames(covs2) <- c("Male",
                     "White", 
                     "Hisp",
                     "Officer",
                    "Street",
                    "Proactive",
                    "Nyears",
                    "Gun",
                    "Taser",
                    "Force",
                    "Nops",
                    "Eig",
                    "outdegree",
                    "indegree",
                    "reciprocation" ,
                    "instar",
                    "outstar",
                    "twopath",
                    "in3star",
                    "out3star",
                    "transitive")


head(covs2)
```

---

# Estimate ALAAMs
Run base and contagion models using the ALAAM routines. (Evaluation disabled by default.)


Set outcome as y_vector
```{r}
y_vector <- invite$grp_invite
```

Run prelimary Markov model that does not include a contagion effect
```{r firstrun}
res.0 <- BayesALAAM(y = y_vector,            # dependent variable
                    ADJ = adj2,              # network
                    covariates = out.degree, # covariates
                    directed = TRUE,         # directed network
                    Iterations = 1000,       # number of iterations
                    saveFreq = 100,          # print and save frequency
                    contagion = 'none')      # type of contagion
```

```{r tracefirstrun}
plot(ts(res.0$Theta[,1:10]))
```

Run a model that that includes a contagion effect (invite) - model 1
```{r}
res.1 <- BayesALAAM(y = y_vector,                                                  # dependent variable
                    ADJ = adj2,                                                    # network
                    covariates = covs2[,c(1,2,3,4,5,6,7,8,9,10,11,13,14,15,21)],   # covariates
                    directed = TRUE,                                               # directed network
                    Iterations = 20000,                                            # number of iterations
                    saveFreq = 200,                                                # print and save frequency
                    contagion ='simple')                                           # contagion ='simple' 


res.1.table <- write.res.table(burnin=10000,                             # should be set sufficiently high
                datamat=res.1$Thetas,                                    # the result from BayesALAAM
                thin=1,                                                  # should be set so that SACF is sufficiently low, impt for CI
                tabname=NULL)                                            # the name appended to the table that is saved

sim.1 <- get.gof.distribution(NumIterations=20000,                       # number of vectors to draw
	                              res=res.1,                               # the ALAAM estimation object contains model/results
	                              burnin=1000,                             # no. iterations discarded from GOF distribution
	                              thinning = 20,                           # no. iterations between sample points
	                              contagion = 'simple')                    # should be the same as for model fitted
	
gof.table(obs.stats=	sim.1$stats,                                       # observed statistics included  not fitted statistics
          sim.stats=	sim.1$Sav.gof,                                     # simulated goodness-of-fit statistics
          name.vec=	sim.1$gof.stats.names,                               # names of statistics calculated
          tabname='ALAAMGofalt',                                         # name of file saved
          pvalues=TRUE,                                                  # posterior predictive p-values
          save.tab ='csv',                                               # save a csv file or a LaTex file
          directed=TRUE,
          Imp.gof = sim.1$Imp.gof)                                       # NB: we have missing values so we need to add these
```

# Assess Convergence and Posterior Summaries
Trace plots, effective sample sizes, and summary tables.

Check how well the MCMC mixes do a traceplot of the poster
```{r trace}
plot(ts(res.1$Thetas))
```

```{r propvar}
Propsigma2 <- cov(res.1$Thetas)
```

---

# Re-estimate with Optimized Proposal Covariance
Use `Propsigma2` to improve mixing.

```{r}
res.2 <- BayesALAAM(y = y_vector,                                                     # dependent variable
                    ADJ = adj2,                                                       # network
                    covariates = covs2[,c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,21)],   # covariates
                    directed = TRUE,                                                  # directed network
                    Iterations = 20000,                                               # number of iterations
                    saveFreq = 200,                                                   # print and save frequency
                    PropSigma = Propsigma2,
                    contagion ='simple')                                              # contagion ='simple' 


res.2.table <- write.res.table(burnin=10000,                             # should be set sufficiently high
                datamat=res.2$Thetas,                                    # the result from BayesALAAM
                thin=1,                                                  # should be set so that SACF is sufficiently low, impt for CI
                tabname=NULL)                                            # the name appended to the table that is saved

sim.2 <- get.gof.distribution(NumIterations=20000,                       # number of vectors to draw
	                              res=res.2,                               # the ALAAM estimation object that contains model and results
	                              burnin=1000,                             # no. iterations discarded from GOF distribution
	                              thinning = 20,                           # no. iterations between sample points
	                              contagion = 'simple')                    # should be the same as for model fitted
	
gof.table(obs.stats=	sim.2$stats,                                       # observed statistics included  not fitted statistics
          sim.stats=	sim.2$Sav.gof,                                     # simulated goodness-of-fit statistics
          name.vec=	sim.2$gof.stats.names,                               # names of statistics calculated
          tabname='ALAAMGofalt',                                         # name of file saved
          pvalues=TRUE,                                                  # posterior predictive p-values
          save.tab ='csv',                                               # save a csv file or a LaTex file
          directed=TRUE,
          Imp.gof = sim.2$Imp.gof)                                       # NB: we have missing values so we need to add these
```

---

# Check Performance and Posteriors
Use ``r 'plotPost'`` to simultaneously plot the posterior distributions, the (serial) autocorrelations, and the trace plots
```{r thirdrunplot}
plotPost(ALAAMresult=res.2,figname='simplecontagion')
```

Produce a results table
```{r restabb2}
write.res.table(burnin=1, # should be set sufficiently high
                datamat=res.2$Thetas, # the result from BayesALAAM
                thin=1, # should be set so that SACF is sufficiently low, important for CI
                tabname=NULL) # the name appended to the table that is saved
```


Calculate likelihoods
```{r indeplike}
logit.est <- glm(res.2$ALAAMobj$y~res.1$ALAAMobj$covariates, family = binomial(link = "logit"))
p <- dim(res.1$Thetas)[2]
thetaRef <- matrix(0,p,1)
thetaRef[1] <- summary(logit.est)$coef[1,1]
thetaRef[3:p] <- summary(logit.est)$coef[2:(p-1),1]

```

Evaluate the deviance across posterior draws
```{r postedev1}
relLike <- aitkinPostDev(ALAAMresult=res.2,# the ALAAM results object
                           burnin=20, # number of parameter draws to be discarded - should eliminate dependence on initial conditions
                           thinning=15, # model selection is more sensitive to serial autocorrelation than point estimates and standard deviations
                           numYsamps=100, # number of simulated vectors to base Metropolis expectation on
                           thetaRef=t(thetaRef), # input the parameters used for reference for evaluating independent likelihood
                           numbridges=20, # 20 bridges should be enough but more will give higher precision
                           Yburnin=1000)

```



